---
title: 'NR States'
author: "Serena Leung"
date: "2/6/2022"
output: html_document
---

NR States (13 states in donor pool) and no covariates

Iter state Function only for NR States
```{r}
# Create generic function that goes through dataprep, synth.out, and plot 
# for state id 2-51

iter_state <- function(dataset, id){   # special cases for id = 1 and 52
  
  if (id == 1) {  
     dataprep.out <- dataprep(dataset, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c((id+1):14),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
     
  } else if (id == 14) {
    
     dataprep.out <- dataprep(dataset, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c(1:(id-1)),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
     
  } else
     dataprep.out <- dataprep(dataset, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c(1:(id-1), (id+1):14),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
  
  synth.out = synth(data.prep.obj = dataprep.out, method = "BFGS")
  
  gaps <- dataprep.out$Y1plot - (dataprep.out$Y0plot %*% synth.out$solution.w)

  gaps <- data.frame("Year"=rownames(gaps), gaps)
  
  return(gaps)

}

```


//////////////////////////////////////// PROP CRIME ////////////////////////////////////////////////////


Read data 
```{r}
library(readxl)
NR_prop_crime <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Property Crime(NR)", skip = 1)

NR_prop_crime = rename(NR_prop_crime, State = "No Raise States")

head(NR_prop_crime)
```


```{r}
NR_prop_crime <- pivot_longer(NR_prop_crime, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

NR_prop_crime

```

```{r}
dataprep.prop <- dataprep(NR_prop_crime, 
         time.predictors.prior = c(2000:2010),
         special.predictors = list(
         list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",        
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = c(2000:2010), # pretreatment period
         time.plot = c(2000:2019)) # time period to be plotted
```


```{r}
synth.prop = synth(data.prep.obj = dataprep.prop, method = "All")
```



```{r}
gaps <- dataprep.prop$Y1plot - (dataprep.prop$Y0plot %*% synth.prop$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)



plot(gaps, type="l", lwd = 2, ylim = c(-500, 500), Ylab = "Gap in Property Crime rate per 100k people")

abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (id in seq(1,14)) {
  state_gap <- iter_state(NR_prop_crime, id)

  lines(state_gap[,1], state_gap[,2], col ="grey")
}
```



Path.plot with California as the treatment
```{r}
path.plot(synth.res=synth.nr, dataprep.res = dataprep.nr, 
          Ylab="Property Crime rate per 100k people", Xlab="year",
          Main = "Property Crime in California", 
          Legend = c("California","Synthetic California"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```

Gap plot California treatment
```{r}
gaps.plot(synth.res = synth.nr, dataprep.res = c(dataprep.nr), Ylab = "Gap in Property Crime rate per 100k people", Xlab= "year", Main = NA)
abline(v = 2014.9166, col = "red")
```

////////////////////////////////////////////////////// LARCENY /////////////////////////////////////////////////////////////////




```{r}
library(readxl)
NR_larc <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Larceny-theft (NR)", skip = 1)

NR_larc = rename(NR_larc, State = "No Raise States")

head(NR_larc)

```



```{r}
NR_larc <- pivot_longer(NR_larc, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

NR_larc

```


Dataprep for No Raise States
```{r}
dataprep.larc <- dataprep(NR_larc, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
```


```{r}
synth.larc = synth(data.prep.obj = dataprep.larc, method = "BFGS")
```




Iterating through the No Raise States 
```{r}
gaps <- dataprep.larc$Y1plot - (dataprep.larc$Y0plot %*% synth.larc$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)



plot(gaps, type="l", lwd = 2, ylim = c(-500, 500), Ylab = "Gap in Larceny Crime rate per 100k people")

abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (id in seq(1,14)) {
  state_gap <- iter_state(NR_larc, id)

  lines(state_gap[,1], state_gap[,2], col ="grey")
}
```

///////////////////////////////////////////////// MOTOR VEHICLE THEFT /////////////////////////////////////////////////


```{r}
library(readxl)
NR_mvt <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Motor Vehicle Theft (NR)", skip = 1)

NR_mvt = rename(NR_mvt, State = "No Raise States")

head(NR_mvt)

```



```{r}
NR_mvt <- pivot_longer(NR_mvt, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

NR_mvt

```



```{r}
dataprep.mvt <- dataprep(NR_mvt, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
```


```{r}
synth.mvt = synth(data.prep.obj = dataprep.mvt, method = "BFGS")
```



Iterating through the No Raise States 
```{r}
gaps <- dataprep.mvt$Y1plot - (dataprep.mvt$Y0plot %*% synth.mvt$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)



plot(gaps, type="l", lwd = 2, ylim = c(-500, 500), Ylab = "Gap in Motor Vehicle Theft rate per 100k people")

abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (id in seq(1,14)) {
  state_gap <- iter_state(NR_mvt, id)

  lines(state_gap[,1], state_gap[,2], col ="grey")
}
```

///////////////////////////////////////////////////////////// ROBBERY ///////////////////////////////////////


```{r}
library(readxl)
NR_rob <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Robbery (NR)", skip = 1)

head(NR_rob)

```



```{r}
NR_rob <- pivot_longer(NR_rob, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

NR_rob

```



```{r}
dataprep.rob <- dataprep(NR_rob, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
```


```{r}
synth.rob = synth(data.prep.obj = dataprep.rob, method = "BFGS")
```



Iterating through the No Raise States 
```{r}
gaps <- dataprep.rob$Y1plot - (dataprep.rob$Y0plot %*% synth.rob$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)



plot(gaps, type="l", lwd = 2, col = "blue", ylim = c(-500, 500), Ylab = "Gap in roblary Crime rate per 100k people")

abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (id in seq(1,14)) {
  state_gap <- iter_state(NR_rob,id)

  lines(state_gap[,1], state_gap[,2], col ="grey")
}
```

//////////////////////////////////////////////////// BURGLARY /////////////////////////////////////////



```{r}
library(readxl)
NR_burg <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Burglary (NR)", skip = 1)

NR_burg = rename(NR_burg, State = "No Raise State")

head(NR_burg)

```



```{r}
NR_burg <- pivot_longer(NR_burg, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

NR_burg

```



```{r}
dataprep.burg <- dataprep(NR_burg, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
```


```{r}
synth.burg = synth(data.prep.obj = dataprep.burg, method = "BFGS")
```



Iterating through the No Raise States 
```{r}
gaps <- dataprep.burg$Y1plot - (dataprep.burg$Y0plot %*% synth.burg$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)



plot(gaps, type="l",  col = "blue" , lwd = 2, ylim = c(-500, 500), Ylab = "Gap in Burglary Crime rate per 100k people")

abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (id in seq(1,14)) {
  state_gap <- iter_state(NR_burg,id)

  lines(state_gap[,1], state_gap[,2], col ="grey")
}
```



/////////////////////////////////////////////// VIOLENT CRIME //////////////////////////////////////////

```{r}
library(readxl)
NR_vio_crime <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Violent Crime (NR)", skip = 1)

NR_vio_crime = rename(NR_vio_crime, State = "No Raise States")

head(NR_vio_crime)

```



```{r}
NR_vio_crime <- pivot_longer(NR_vio_crime, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

NR_vio_crime

```



```{r}
dataprep.vio <- dataprep(NR_vio_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
```


```{r}
synth.vio = synth(data.prep.obj = dataprep.vio, method = "BFGS")
```



Iterating through the No Raise States 
```{r}
gaps <- dataprep.vio$Y1plot - (dataprep.vio$Y0plot %*% synth.vio$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)



plot(gaps, type="l", col = "blue", lwd = 2, ylim = c(-500, 500), Ylab = "Gap in vio_crimelary Crime rate per 100k people")

abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (id in seq(1,14)) {
  state_gap <- iter_state(NR_vio_crime, id)

  lines(state_gap[,1], state_gap[,2], col ="grey")
}
```


NR States (No Covariates) Weights
```{r}

cov_weights <- data.frame(synth.prop$solution.w, synth.larc$solution.w, synth.mvt$solution.w, synth.burg$solution.w, synth.rob$solution.w, synth.vio$solution.w)

names(cov_weights) <- c("Property Crime", "Larceny", "Motor Vehicle Theft", "Burglary", "Robbery", "Violent Crime")

cov_weights$ID <- rownames(cov_weights)

pred_prop %>% select("ID", "State") %>% 
  distinct() %>% 
  mutate("ID" = as.character(ID))%>% 
  right_join(cov_weights) %>% 
  mutate_if(is.numeric, round, digits = 3)



```




































