---
title: "Synth"
author: "Serena Leung"
date: "11/23/2021"
output: html_document
---

```{r}
library(Synth)
library(tidyverse)
```


Read data 
```{r}
library(readxl)
prop_crime <- read_excel("~/Thesis/Master State Crime.xlsx", 
    sheet = "Property Crime", skip = 1)
prop_crime
```


```{r}
prop_crime <- pivot_longer(prop_crime, cols = c(-State), names_to = "year", values_to = "rate") %>% 
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

prop_crime


```


```{r}
dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       # ????? 
         unit.names.variable = "State", # ?????
         time.variable = "year", 
         treatment.identifier = 5, 
         controls.identifier = c(1:4, 6:52),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
```




convert to long format 
(state, year, rate)

Predictors: 
X1 = CA 2000 - 2010 before treatment (11 x 1) 
  - since looking at 11 yr gap
  
X0 = 51 states 2000 - 2010 before treatment (11 x 51)

Outcomes:
Z1 = CA 2014 - 2019 after treatment (6 x 1)
Z0 = 51 states after treatment (6 x 51)


```{r}
synth.out = synth(data.prep.obj = dataprep.out, method = "BFGS")
```



Summary Tables
```{r}
synth.tables = synth.tab(dataprep.res = dataprep.out, synth.res = synth.out)

names(synth.tables)
```

Table comparing pre-treatment predictor values for the treated unit, the synthetic control, and all the units in the sample (What does this mean in our case?)
```{r}
synth.tables$tab.pred
```


```{r}
synth.tables$tab.loss
```

Weights for each state
```{r}
synth.tables$tab.w
```


Plotting before and after treatment
```{r}
path.plot(synth.res=synth.out, dataprep.res = dataprep.out, 
          Ylab="Property Crime rate per 100k people", Xlab="year",
          Legend = c("California","Synthetic California"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```
Gaps 
```{r}
gaps.plot(synth.res = synth.out, dataprep.res = c(dataprep.out), Ylab = "Gap in Property Crime rate per 100k people", Xlab= "year", Main = NA)
abline(v = 2014.9166, col = "red")
```





Adjusting pretreatment periods

```{r}
prop_crime <- as.data.frame(prop_crime)
```


```{r}
dataprep2.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2005,
         special.predictors = list(
         list("rate", 2000:2005, "mean")),
         dependent = "rate", 
         unit.variable = "ID",        
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 5, 
         controls.identifier = c(1:4, 6:52),
         time.optimize.ssr = 2000:2005, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
```

```{r}
synth2.out = synth(data.prep.obj = dataprep2.out, method = "BFGS")
```



```{r}
path.plot(synth.res=synth2.out, dataprep.res = dataprep2.out, 
          Ylab="Property Crime rate per 100k people", Xlab="year",
          Legend = c("California","Synthetic California"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```


```{r}
iter_year <- function(year){
  
   dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:year,
         special.predictors = list(
          list("rate", 2000:year, "mean")),  # am I supposed to change this?
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 5, 
         controls.identifier = c(1:4, 6:52),
         time.optimize.ssr = 2000:year, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
   
  synth.out = synth(data.prep.obj = dataprep.out, method = "BFGS")
  
  path.plot(synth.res=synth.out, dataprep.res = dataprep.out, 
          Ylab="Property Crime rate per 100k people", Xlab="year",
          Legend = c(paste("California", year),"Synthetic"),
          Legend.position = "bottomleft", tr.intake = 2014.9166)
  abline(v = 2014.9166, col = "red")
  rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
  
}
```


```{r}

for(i in seq(2000,2010)){
  iter_year(i)
}

```







Leave California out, use Alabama as treatment state, remaining 50 states are in the control group

Iterate over different states as the treatment group

```{r}
# Create generic function that goes through dataprep, synth.out, and plot 
# for state id 2-51

iter_state <- function(id){   # special cases for id = 1 and 52
  
  if (id == 1) {  
     dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c((id+1):52),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
     
  } else if (id == 52) {
    
     dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c(1:(id-1)),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
     
  } else
     dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c(1:(id-1), (id+1):52),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
  
  synth.out = synth(data.prep.obj = dataprep.out, method = "BFGS")
  
  path.plot(synth.res=synth.out, dataprep.res = dataprep.out, 
          Ylab="Property Crime rate per 100k people", Xlab="year",
          Legend = c(id,"Synthetic"),
          Legend.position = "bottomleft", tr.intake = 2014.9166)
  abline(v = 2014.9166, col = "red")
  rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
}

```




```{r}
for (id in seq(1,52)) {
  iter_state(id)
}


```




```{r}
dataprep3.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       # ????? 
         unit.names.variable = "State", # ?????
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:4, 6:52),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
```

```{r}
synth.out = synth(data.prep.obj = dataprep3.out, method = "BFGS")
```

```{r}
path.plot(synth.res=synth.out, dataprep.res = dataprep3.out, 
          Ylab="Property Crime rate per 100k people", Xlab="year",
          Legend = c("Alabama","Synthetic Alabama"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```

```{r}
gaps.plot(synth.res = synth.out, dataprep.res = dataprep3.out, Ylab = "Gap in Property Crime rate per 100k people", Xlab= "year", Main = NA)
abline(v = 2014.9166, col = "red")
```



