---
title: "Synth"
author: "Serena Leung"
date: "11/23/2021"
output: html_document
---

```{r}
library(Synth)
library(tidyverse)
```


Read data 
```{r}
library(readxl)
prop_crime <- read_excel("~/Thesis/Master State Crime.xlsx", 
    sheet = "Property Crime", skip = 1)
prop_crime
```


```{r}
prop_crime <- pivot_longer(prop_crime, cols = c(-State), names_to = "year", values_to = "rate") %>% 
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

prop_crime


```


```{r}
dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       # ????? 
         unit.names.variable = "State", # ?????
         time.variable = "year", 
         treatment.identifier = 5, 
         controls.identifier = c(1:4, 6:52),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
```




convert to long format 
(state, year, rate)

Predictors: 
X1 = CA 2000 - 2010 before treatment (11 x 1) 
  - since looking at 11 yr gap
  
X0 = 51 states 2000 - 2010 before treatment (11 x 51)

Outcomes:
Z1 = CA 2014 - 2019 after treatment (6 x 1)
Z0 = 51 states after treatment (6 x 51)


```{r}
synth.out = synth(data.prep.obj = dataprep.out, method = "BFGS")
```






Manually calculating gaps

```{r}
# Create generic function that goes through dataprep, synth.out, and plot 
# for state id 2-51

iter_state <- function(id){   # special cases for id = 1 and 52
  
  if (id == 1) {  
     dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c((id+1):52),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
     
  } else if (id == 52) {
    
     dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c(1:(id-1)),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
     
  } else
     dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c(1:(id-1), (id+1):52),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
  
  synth.out = synth(data.prep.obj = dataprep.out, method = "BFGS")
  
  gaps <- dataprep.out$Y1plot - (dataprep.out$Y0plot %*% synth.out$solution.w)

  gaps <- data.frame("Year"=rownames(gaps), gaps)
  
  return(gaps)
  
  # path.plot(synth.res=synth.out, dataprep.res = dataprep.out, 
  #         Ylab="Property Crime rate per 100k people", Xlab="year",
  #         Legend = c(id,"Synthetic"),
  #         Legend.position = "bottomleft", tr.intake = 2014.9166)
  # abline(v = 2014.9166, col = "red")
  # rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
}

```


```{r}
gaps <- dataprep.out$Y1plot - (dataprep.out$Y0plot %*% synth.out$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)


AL_gap <- iter_state(1)
AK_gap <- iter_state(2)


plot(gaps, type="l", lwd = 2, ylim = c(-500, 500), Ylab = "Gap in Property Crime rate per 100k people")
# lines(AL_gap$Year, AL_gap$X1, col ="grey")
# lines(AK_gap$Year, AK_gap[,2], col ="grey")
# 
# AK_gap[,2]
abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)
for (id in seq(1,52)) {
  state_gap <- iter_state(id)

  lines(state_gap[,1], state_gap[,2], col ="grey")
}

```

- How do I make this graph less messy? What criteria do I use to choose which tests I should use?
- Why is there a peak at 2010? 







///////////////////////////// Adjusting pretreatment periods//////////////////////////////

```{r}
prop_crime <- as.data.frame(prop_crime)
```


```{r}
dataprep2.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2005,
         special.predictors = list(
         list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",        
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 5, 
         controls.identifier = c(1:4, 6:52),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
```

```{r}
synth2.out = synth(data.prep.obj = dataprep2.out, method = "BFGS")
```



```{r}
path.plot(synth.res=synth2.out, dataprep.res = dataprep2.out, 
          Ylab="Property Crime rate per 100k people", Xlab="year",
          Legend = c("California","Synthetic California"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```


```{r}
iter_year <- function(year){
  
   dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:year,
         special.predictors = list(
          list("rate", 2000:year, "mean")),  # am I supposed to change this?
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 5, 
         controls.identifier = c(1:4, 6:52),
         time.optimize.ssr = 2000:year, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
   
  synth.out = synth(data.prep.obj = dataprep.out, method = "BFGS")
  
  gaps <- dataprep.out$Y1plot - (dataprep.out$Y0plot %*% synth.out$solution.w)

  gaps <- data.frame("Year"=rownames(gaps), gaps)
  
  return(gaps)
  
  # path.plot(synth.res=synth.out, dataprep.res = dataprep.out, 
  #         Ylab="Property Crime rate per 100k people", Xlab="year",
  #         Legend = c(paste("California", year),"Synthetic"),
  #         Legend.position = "bottomleft", tr.intake = 2014.9166)
  # abline(v = 2014.9166, col = "red")
  # rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
  
}
```


```{r}

gaps <- dataprep2.out$Y1plot - (dataprep2.out$Y0plot %*% synth2.out$solution.w)

gaps <- data.frame("Year"=rownames(gaps), gaps)


plot(gaps, type="l", lwd = 6, ylim = c(-500, 500))
abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (i in seq(2000,2010)) {
  year_gap <- iter_year(i)

  lines(year_gap[,1], year_gap[,2], col ="grey")
}

# for(i in seq(2000,2010)){
#   iter_year(i)
# }

```







Leave California out, use Alabama as treatment state, remaining 50 states are in the control group

Iterate over different states as the treatment group

```{r}
# Create generic function that goes through dataprep, synth.out, and plot 
# for state id 2-51

old_iter_state <- function(id){   # special cases for id = 1 and 52
  
  if (id == 1) {  
     dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c((id+1):52),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
     
  } else if (id == 52) {
    
     dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c(1:(id-1)),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
     
  } else
     dataprep.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c(1:(id-1), (id+1):52),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
  
  synth.out = synth(data.prep.obj = dataprep.out, method = "BFGS")
  
  path.plot(synth.res=synth.out, dataprep.res = dataprep.out, 
          Ylab="Property Crime rate per 100k people", Xlab="year",
          Legend = c(id,"Synthetic"),
          Legend.position = "bottomleft", tr.intake = 2014.9166)
  abline(v = 2014.9166, col = "red")
  rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
}

```




```{r}
for (id in seq(1,52)) {
  iter_state(id)
}

```




```{r}
dataprep3.out <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       # ????? 
         unit.names.variable = "State", # ?????
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:4, 6:52),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
```

```{r}
synth.out = synth(data.prep.obj = dataprep3.out, method = "BFGS")
```

```{r}
path.plot(synth.res=synth.out, dataprep.res = dataprep3.out, 
          Ylab="Property Crime rate per 100k people", Xlab="year",
          Legend = c("Alabama","Synthetic Alabama"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```

```{r}
gaps.plot(synth.res = synth.out, dataprep.res = dataprep3.out, Ylab = "Gap in Property Crime rate per 100k people", Xlab= "year", Main = NA)
abline(v = 2014.9166, col = "red")
```

///////////////////////////////////SC WITH NO RAISE STATES/////////////////////////////////////////////////


Read data 
```{r}
library(readxl)
NR_prop_crime <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Property Crime(NR)", skip = 1)

NR_prop_crime = rename(NR_prop_crime, State = "No Raise States")

head(NR_prop_crime)
```


```{r}
NR_prop_crime <- pivot_longer(NR_prop_crime, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

# NR_prop_crime$`No Raise States`
# # 
# as.numeric(factor(NR_prop_crime$`No Raise States`))

NR_prop_crime

```


```{r}
dataprep.nr <- dataprep(prop_crime, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       # ????? 
         unit.names.variable = "State", # ?????
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
```


```{r}
synth.nr = synth(data.prep.obj = dataprep.nr, method = "BFGS")
```



```{r}
gaps <- dataprep.nr$Y1plot - (dataprep.nr$Y0plot %*% synth.nr$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)


AL_gap <- iter_state(1)
AK_gap <- iter_state(2)


plot(gaps, type="l", lwd = 2, ylim = c(-500, 500), Ylab = "Gap in Property Crime rate per 100k people")
# lines(AL_gap$Year, AL_gap$X1, col ="grey")
# lines(AK_gap$Year, AK_gap[,2], col ="grey")
# 
# AK_gap[,2]
abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)
for (id in seq(1,14)) {
  state_gap <- iter_state(id)

  lines(state_gap[,1], state_gap[,2], col ="grey")
}
```



Path.plot with California as the treatment
```{r}
path.plot(synth.res=synth.nr, dataprep.res = dataprep.nr, 
          Ylab="Property Crime rate per 100k people", Xlab="year",
          Legend = c("California","Synthetic California"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```

Gap plot California treatment
```{r}
gaps.plot(synth.res = synth.nr, dataprep.res = c(dataprep.nr), Ylab = "Gap in Property Crime rate per 100k people", Xlab= "year", Main = NA)
abline(v = 2014.9166, col = "red")
```
