---
title: "Predictor Variables"
author: "Serena Leung"
date: "1/30/2022"
output: html_document
---


## Loading Predictor Variables 
```{r}
library(Synth)
library(readxl)
library(tidyverse)

# Unemployment Rate
UR_Data <- read_excel("~/Thesis/Predictors/UR Data.xlsx", 
    sheet = "Sheet1")

UR_Data <- UR_Data %>% filter(Year >= 2000) %>% mutate(year = as.numeric(Year)) %>% select(-c("FIPS Code", Total, Month, Year)) 
```

```{r}
student_exp <- read_excel("~/Thesis/Predictors/Student Expenditure 08.xls") %>% 
  pivot_longer(cols = c(-State), names_to = "year", values_to = "pupil_exp") %>% 
  mutate(year = as.numeric(year))
head(student_exp)
```

State GDP
```{r}
library(readr)
state_gdp <- read_csv("~/Thesis/Predictors/GDP by State.csv", skip = 4) %>% 
  select(-c(GeoFips, "1997", "1998", "1999")) %>% 
  pivot_longer(cols = c(-GeoName), names_to = "year", values_to = "state_gdp") %>% 
  rename(State = GeoName) %>% 
  mutate(year = as.numeric(year))

  
head(state_gdp)
```


Iter state function for plotting gap plot 
(Specifically for No Raise States)

```{r}
# Create generic function that goes through dataprep, synth.out, and plot 
# for state id 2-51

iter_state <- function(dataset, id){   # special cases for id = 1 and 52
  
  if (id == 1) {  
     dataprep.out <- dataprep(dataset, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c((id+1):14),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
     
  } else if (id == 14) {
    
     dataprep.out <- dataprep(dataset, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c(1:(id-1)),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
     
  } else
     dataprep.out <- dataprep(dataset, 
         time.predictors.prior = 2000:2010,
         special.predictors = list(
          list("rate", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",       
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = id, 
         controls.identifier = c(1:(id-1), (id+1):14),
         time.optimize.ssr = 2000:2010, # pretreatment period
         time.plot = 2000:2019) # time period to be plotted
  
  synth.out = synth(data.prep.obj = dataprep.out, method = "BFGS")
  
  gaps <- dataprep.out$Y1plot - (dataprep.out$Y0plot %*% synth.out$solution.w)

  gaps <- data.frame("Year"=rownames(gaps), gaps)
  
  return(gaps)

}

```


## Property Crime (No Raise States)
```{r}
library(readxl)
NR_prop_crime <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Property Crime(NR)", skip = 1)

NR_prop_crime = rename(NR_prop_crime, State = "No Raise States")

head(NR_prop_crime)
```


```{r}
pred_prop <- pivot_longer(NR_prop_crime, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))


pred_prop

```



Join tables together
```{r}
pred_prop <- pred_prop %>% 
  left_join(UR_Data) %>% 
  left_join(student_exp) %>% 
  left_join(state_gdp)

pred_prop
```

Calculating RMSE
```{r}
y0 <- pred_prop %>% filter(State != "California") %>% select(State, year, rate) %>% pivot_wider(names_from = year, values_from = rate) %>% as.matrix()

y0 <- y0[,2:ncol(y0)] %>% as.matrix()

y0 <- apply(y0, 2,as.numeric)

y1 <- pred_prop %>% filter(State == "California") %>% select(State, year, rate) %>% pivot_wider(names_from = year, values_from = rate)
y1 <- apply(y1[,2:ncol(y1)], 2,as.numeric)

yhat <-t(as.matrix(W))%*%y0

synthres<-y1-yhat

RMSE_pre<-sqrt(sum(synthres[1:11]^2)/11)
RMSE_postall<-sqrt(sum(synthres[12:20]^2)/9)
RMSE_post<-sqrt(sum(synthres[15:20]^2)/5)
```
post should not fit well if there is effect 
pre should fit well bc you want pretreatment to mimic treatment  






```{r}
dataprep.prop <- dataprep(pred_prop,
         predictors = c("UR", "pupil_exp", "state_gdp"),
         predictors.op = "mean",
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean"),
         list("UR", 2000:2010, "mean"),
         list("pupil_exp",2000:2010, "mean"),
         list("state_gdp", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",        
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = c(2000:2010), # pretreatment period
         time.plot = c(2000:2019)) # time period to be plotted
```


```{r}
dataprep.prop$X0

dataprep.prop$X1
```


```{r}
synth.prop = synth(data.prep.obj = dataprep.prop, method = "All")
```

Covariates Weight Table
```{r}

cov_weights <- data.frame(synth.prop$solution.w, synth.larc$solution.w, synth.mvt$solution.w, synth.burg$solution.w, synth.rob$solution.w, synth.vio$solution.w)

names(cov_weights) <- c("Property Crime", "Larceny", "Motor Vehicle Theft", "Burglary", "Robbery", "Violent Crime")

cov_weights$ID <- rownames(cov_weights)

pred_prop %>% select("ID", "State") %>% 
  distinct() %>% 
  mutate("ID" = as.character(ID))%>% 
  right_join(cov_weights) %>% 
  mutate_if(is.numeric, round, digits = 3)



```



```{r}
path.plot(synth.res=synth.prop, dataprep.res = dataprep.prop, 
          Ylab="Property Crime rate per 100k people", Xlab="Year", 
          Main = "Property Crime in California", 
          Legend = c("California","Synthetic California"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```
```{r}
gaps <- dataprep.prop$Y1plot - (dataprep.prop$Y0plot %*% synth.prop$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)


plot(gaps, type="l", lwd = 2, ylim = c(-500, 500), ylab = "Gap in Property Crime rate per 100k people", main = "Gap Plot for Property Crime")

abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (id in seq(1,14)) {
  state_gap <- iter_state(pred_prop, id)
  lines(state_gap[,1], state_gap[,2], col ="grey")
}

```

#### LARCENY ####### 


```{r}
library(readxl)
NR_larc <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Larceny-theft (NR)", skip = 1)

NR_larc = rename(NR_larc, State = "No Raise States")

head(NR_larc)

```



```{r}
NR_larc <- pivot_longer(NR_larc, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

NR_larc

```



```{r}

pred_larc <- NR_larc %>% 
  left_join(UR_Data) %>% 
  left_join(student_exp) %>% 
  left_join(state_gdp)
```



```{r}
dataprep.larc <- dataprep(pred_larc,
         predictors = c("UR", "pupil_exp", "state_gdp"),
         predictors.op = "mean",
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean"),
         list("UR", 2000:2010, "mean"),
         list("pupil_exp",2000:2010, "mean"),
         list("state_gdp", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",        
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = c(2000:2010), # pretreatment period
         time.plot = c(2000:2019)) # time period to be plotted
```


```{r}
synth.larc = synth(data.prep.obj = dataprep.larc, method = "All")
```

```{r}
path.plot(synth.res=synth.larc, dataprep.res = dataprep.larc, 
          Ylab="Property Crime rate per 100k people", Xlab="Year", 
          Main = "Larceny in California", 
          Legend = c("California","Synthetic California"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```

```{r}
gaps <- dataprep.larc$Y1plot - (dataprep.larc$Y0plot %*% synth.larc$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)


plot(gaps, type="l", lwd = 2, ylim = c(-500, 500), ylab = "Gap in Larceny rate per 100k people", main = "Gap Plot for Larceny Theft")

abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (id in seq(1,14)) {
  state_gap <- iter_state(pred_prop, id)
  lines(state_gap[,1], state_gap[,2], col ="grey")
}

```


##### MOTOR VEHICLE THEFT #####

```{r}
library(readxl)
NR_mvt <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Motor Vehicle Theft (NR)", skip = 1)

NR_mvt = rename(NR_mvt, State = "No Raise States")

head(NR_mvt)

```



```{r}
NR_mvt <- pivot_longer(NR_mvt, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

NR_mvt

```


```{r}

pred_mvt <- NR_mvt %>% 
  left_join(UR_Data) %>% 
  left_join(student_exp) %>% 
  left_join(state_gdp)
```



```{r}
dataprep.mvt <- dataprep(pred_mvt,
         predictors = c("UR", "pupil_exp", "state_gdp"),
         predictors.op = "mean",
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean"),
         list("UR", 2000:2010, "mean"),
         list("pupil_exp",2000:2010, "mean"),
         list("state_gdp", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",        
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = c(2000:2010), # pretreatment period
         time.plot = c(2000:2019)) # time period to be plotted
```


```{r}
synth.mvt = synth(data.prep.obj = dataprep.mvt, method = "All")
```

```{r}
path.plot(synth.res=synth.mvt, dataprep.res = dataprep.mvt, 
          Ylab="Motor Vehicle Theft rate per 100k people", Xlab="Year", 
          Main = "Motor Vehicle Theft in California", 
          Legend = c("California","Synthetic California"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```

```{r}
gaps <- dataprep.mvt$Y1plot - (dataprep.mvt$Y0plot %*% synth.mvt$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)


plot(gaps, type="l", lwd = 2, ylim = c(-500, 500), ylab = "Gap in Motor Vehicle Theft rate per 100k people", main = "Gap Plot for Motor Vehicle Theft")

abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (id in seq(1,14)) {
  state_gap <- iter_state(pred_prop, id)
  lines(state_gap[,1], state_gap[,2], col ="grey")
}

```


##### ROBBERY #####


```{r}
library(readxl)
NR_rob <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Robbery (NR)", skip = 1)

head(NR_rob)

```



```{r}
NR_rob <- pivot_longer(NR_rob, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

NR_rob

```


```{r}

pred_rob <- NR_rob %>% 
  left_join(UR_Data) %>% 
  left_join(student_exp) %>% 
  left_join(state_gdp)
```



```{r}
dataprep.rob <- dataprep(pred_rob,
         predictors = c("UR", "pupil_exp", "state_gdp"),
         predictors.op = "mean",
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean"),
         list("UR", 2000:2010, "mean"),
         list("pupil_exp",2000:2010, "mean"),
         list("state_gdp", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",        
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = c(2000:2010), # pretreatment period
         time.plot = c(2000:2019)) # time period to be plotted
```


```{r}
synth.rob = synth(data.prep.obj = dataprep.rob, method = "All")
```

```{r}
path.plot(synth.res=synth.rob, dataprep.res = dataprep.rob, 
          Ylab="Robbery rate per 100k people", Xlab="Year", 
          Main = "Robbery in California", 
          Legend = c("California","Synthetic California"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```

```{r}
gaps <- dataprep.rob$Y1plot - (dataprep.rob$Y0plot %*% synth.rob$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)


plot(gaps, type="l", lwd = 2, ylim = c(-500, 500), ylab = "Gap in Robbery rate per 100k people", main = "Gap Plot for Robbery")

abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (id in seq(1,14)) {
  state_gap <- iter_state(pred_prop, id)
  lines(state_gap[,1], state_gap[,2], col ="grey")
}

```


##### ROBBERY ######

```{r}
library(readxl)
NR_burg <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Burglary (NR)", skip = 1)

NR_burg = rename(NR_burg, State = "No Raise State")

head(NR_burg)

```



```{r}
NR_burg <- pivot_longer(NR_burg, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

NR_burg

```



```{r}

pred_burg <- NR_burg %>% 
  left_join(UR_Data) %>% 
  left_join(student_exp) %>% 
  left_join(state_gdp)
```



```{r}
dataprep.burg <- dataprep(pred_burg,
         predictors = c("UR", "pupil_exp", "state_gdp"),
         predictors.op = "mean",
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean"),
         list("UR", 2000:2010, "mean"),
         list("pupil_exp",2000:2010, "mean"),
         list("state_gdp", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",        
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = c(2000:2010), # pretreatment period
         time.plot = c(2000:2019)) # time period to be plotted
```


```{r}
synth.burg = synth(data.prep.obj = dataprep.burg, method = "All")
```

```{r}
path.plot(synth.res=synth.burg, dataprep.res = dataprep.burg, 
          Ylab="Burglary rate per 100k people", Xlab="Year", 
          Main = "Burglary in California", 
          Legend = c("California","Synthetic California"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```

```{r}
gaps <- dataprep.burg$Y1plot - (dataprep.burg$Y0plot %*% synth.burg$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)


plot(gaps, type="l", lwd = 2, ylim = c(-500, 500), ylab = "Gap in Burglary rate per 100k people", main = "Gap Plot for Burglary")

abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (id in seq(1,14)) {
  state_gap <- iter_state(pred_prop, id)
  lines(state_gap[,1], state_gap[,2], col ="grey")
}

```


##### Violent Crime ##### 

```{r}
library(readxl)
NR_vio_crime <- read_excel("~/Thesis/No Raise Master State Crime.xlsx", 
    sheet = "Violent Crime (NR)", skip = 1)

NR_vio_crime = rename(NR_vio_crime, State = "No Raise States")

head(NR_vio_crime)

```



```{r}
NR_vio_crime <- pivot_longer(NR_vio_crime, cols = c(-State), names_to = "year", values_to = "rate") %>%
  transform(ID = as.numeric(factor(State)), year = as.numeric(year))

NR_vio_crime

```

```{r}

pred_vio <- NR_vio_crime %>% 
  left_join(UR_Data) %>% 
  left_join(student_exp) %>% 
  left_join(state_gdp)
```



```{r}
dataprep.vio <- dataprep(pred_vio,
         predictors = c("UR", "pupil_exp", "state_gdp"),
         predictors.op = "mean",
         time.predictors.prior = 2000:2010,
         special.predictors = list(
         list("rate", 2000:2010, "mean"),
         list("UR", 2000:2010, "mean"),
         list("pupil_exp",2000:2010, "mean"),
         list("state_gdp", 2000:2010, "mean")),
         dependent = "rate", 
         unit.variable = "ID",        
         unit.names.variable = "State", 
         time.variable = "year", 
         treatment.identifier = 1, 
         controls.identifier = c(2:14),
         time.optimize.ssr = c(2000:2010), # pretreatment period
         time.plot = c(2000:2019)) # time period to be plotted
```


```{r}
synth.vio = synth(data.prep.obj = dataprep.vio, method = "All")
```

```{r}
path.plot(synth.res=synth.vio, dataprep.res = dataprep.vio, 
          Ylab="Violent Crime rate per 100k people", Xlab="Year", 
          Main = "Violent Crime in California", 
          Legend = c("California","Synthetic California"),
          Legend.position = "bottomleft")
abline(v = 2014.9166, col = "red")
rect(xleft=2011,xright = 2014 ,ybottom=0,ytop= 5000, density=10, col = "gray") # Realignment
```

```{r}
gaps <- dataprep.vio$Y1plot - (dataprep.vio$Y0plot %*% synth.vio$solution.w)

gaps<- data.frame("Year"=rownames(gaps), gaps)


plot(gaps, type="l", lwd = 2, ylim = c(-500, 500), ylab = "Gap in Violent Crime rate per 100k people", main = "Gap Plot for Violent Crime")

abline(v = 2014.9166, col = "red")
abline(h = 0, lty = 2)

for (id in seq(1,14)) {
  state_gap <- iter_state(pred_prop, id)
  lines(state_gap[,1], state_gap[,2], col ="grey")
}

```








####################### CALCULATING RMSE FOR EACH OUTCOME #################################

Goal: Calculate RMSE for each state for each outcome 
```{r}
me){
  
calc_RMSE <- function(id, outco
  # Control States 
  y0 <- outcome %>% filter(ID != id) %>% select(State, year, rate) %>% pivot_wider(names_from = year,             values_from = rate) %>% as.matrix()
  y0 <- y0[,2:ncol(y0)] %>% as.matrix()
  y0 <- apply(y0, 2,as.numeric)values_from = rate)
  y1 <- apply(y1[,2:ncol(y1)], 2,as.numeric)
  
  # Synthetic State values
  
  # Treated State
  y1 <- outcome %>% filter(ID == id) %>% select(State, year, rate) %>% pivot_wider(names_from = year, 
        
  yhat <-t(as.matrix(W))%*%y0
  
  # True - Synthetic
  synthres<-y1-yhat
  
  RMSE_pre<-sqrt(sum(synthres[1:11]^2)/11)
  RMSE_postall<-sqrt(sum(synthres[12:20]^2)/9)
  RMSE_post<-sqrt(sum(synthres[15:20]^2)/5)
  
  return(RMSE_post/RMSE_pre)
}
```

```{r}
RMSE_loop <- function(outcome){
  RMSE_values <- rep(NA, 14)
  
  for (i in 1:14) {
    RMSE_values[i] <- calc_RMSE(i, outcome)
  }
  return(RMSE_values)
}

prop_RMSE <- RMSE_loop(pred_prop)
larc_RMSE <- RMSE_loop(pred_larc)
mvt_RMSE <- RMSE_loop(pred_mvt)
rob_RMSE <- RMSE_loop(pred_rob)
burg_RMSE <- RMSE_loop(pred_burg)
vio_RMSE <- RMSE_loop(pred_vio)

```

```{r}
# Create dataframe that holds RMSE values for each outcome 
inter_df <- data.frame(ID = 1:14, prop_RMSE, larc_RMSE, mvt_RMSE, rob_RMSE, burg_RMSE, vio_RMSE)

# need states 

RMSE_df <- pred_prop %>% select(State, ID) %>% distinct() %>% left_join(inter_df)
colnames(RMSE_df) <- c("State", "ID", "Property Crime", "Larceny Theft", 
                       "Motor Vehicle Theft", "Robbery", "Burglary", "Violent Crime")

RMSE_long <- RMSE_df %>% pivot_longer(-c(State, ID), names_to = "outcome", values_to = "values")

library(ggplot2)
library(forcats)
library(tidytext)



ggplot(RMSE_long, aes(values, reorder_within(State, values, outcome), fill = outcome)) +
     geom_bar(stat = 'identity', show.legend = FALSE) +
     scale_y_reordered() +
     facet_wrap(.~outcome, ncol=3,scales = "free_y") +
    xlab("RMSE Ratio") + ylab("States")
    

```




